Towards a Mechanized Proof of Noninterference

### 2013-12-19 paper/paper.tex
Typos.

The last decade has also seen a surge of research in languages which inforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.  These languages provide features which programmers can use to enforce integrity and confidentiality of data as it is used throughout computations.

@@

The language's type system includes subtyping, which allows data that have a
lower label to flow into higher labelled data.  We use the notatoin $x \leq y$
to denote that $x$ is a subtype of $y$, whether both $x$ and $y$ are labels or
types (as long as both are the same kind of value).  The rules for subtyping are
given in Figure~\ref{fig:subtype_rules}.  Rule \textsc{S-Int} states that
integers are subtypes of integers with equivalent or higher labels in the label
lattice.  Similarly, rule \textsc{S-Abs} gives the standard subtyping
relationship for functions along with the same label relation.

@@

\textsc{V-Bracket} and \textsc{E-Bracket} both require that there exists a label
from the set $H$ which representes an upward closed sublattice of the labels in
the system representing what is considered to be ``high'' security for the
purposes of the proof of noninterference, discussed later.  These rules require
that the types of each branch of a bracketed value or expression can be typed
with a high security label.  Furthermore, \textsc{E-Bracket} requires that the
$pc$ of the subexpressions must be at least as high as this same label,
requiring that we treat everything in the subexpression as affected by high
security information.

@@

  As a result, we know that $e$ reduces to an answer, because otherwise it would
  be stuck and this would mean that one of the projections is also stuck by
  Lemma~\ref{lem:stuck_bracket}, a contradicton.  Thus, we know that if both
  projections step to an answer, so must the original expression.
\end{proof}

@@

  \begin{case}[lift-if]
    $e = \text{if } <v_1 | v_2> \: e_1 \: e_2$ and $e' = e_i$ for some $i \in
    \{1, 2\}$ in this case.  Applying Lemma~\ref{lem:guard} to the first premise
    of \textsc{E-If} gives $l' \in H$ and $l' \guards t$.  Applying
    Lemma~\ref{lem:projection} to the intial premises of \textsc{E-If} and using
    the result and applying \textsc{E-If} gives $\typeRule{pc \sqcup l}{e_i}$.
    We can then apply \textsc{E-Bracket} to get the goal.
  \end{case}

@@

The noninteference theorem shows that if we have an expression that is typed
with a low label, then replacing any bracketed value in the expression with
either projection will result in the same value for the overall expression.  In
other words, if we vary any high security item in a low security expression, the
value must not be affected.

@@

We found that the process of mechanizing the lemmas that we proved was a useful
exercise in checking our understanding of our language, statements, and the work
of Simonet and Pottier which we based our approach off of.  We found ourselves
often surprised by the number of items that were elided by the proofs in the
paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
natural languae but turned out to be 40 lines of Coq code, not including all of
the supporting definitions and rules.

@@

Once a full Coq proof of noninterference exists, a next step would be to
integrate this proof and all of the related definitions into a verified type
checking program, which could be intergrated into a verified compiler.  We
already know that a full verified compiler has been done for C, and this gives
us hope that it would be possible to implement something similar for a language
like Jif.

>>>>

The last decade has also seen a surge of research in languages which enforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.  These languages provide features which programmers can use to enforce integrity and confidentiality of data as it is used throughout computations.

@@

The language's type system includes subtyping, which allows data that have a
lower label to flow into higher labelled data.  We use the notation $x \leq y$
to denote that $x$ is a subtype of $y$, whether both $x$ and $y$ are labels or
types (as long as both are the same kind of value).  The rules for subtyping are
given in Figure~\ref{fig:subtype_rules}.  Rule \textsc{S-Int} states that
integers are subtypes of integers with equivalent or higher labels in the label
lattice.  Similarly, rule \textsc{S-Abs} gives the standard subtyping
relationship for functions along with the same label relation.

@@

\textsc{V-Bracket} and \textsc{E-Bracket} both require that there exists a label
from the set $H$ which represents an upward closed sublattice of the labels in
the system representing what is considered to be ``high'' security for the
purposes of the proof of noninterference, discussed later.  These rules require
that the types of each branch of a bracketed value or expression can be typed
with a high security label.  Furthermore, \textsc{E-Bracket} requires that the
$pc$ of the subexpressions must be at least as high as this same label,
requiring that we treat everything in the subexpression as affected by high
security information.

@@

  As a result, we know that $e$ reduces to an answer, because otherwise it would
  be stuck and this would mean that one of the projections is also stuck by
  Lemma~\ref{lem:stuck_bracket}, a contradiction.  Thus, we know that if both
  projections step to an answer, so must the original expression.
\end{proof}

@@

  \begin{case}[lift-if]
    $e = \text{if } <v_1 | v_2> \: e_1 \: e_2$ and $e' = e_i$ for some $i \in
    \{1, 2\}$ in this case.  Applying Lemma~\ref{lem:guard} to the first premise
    of \textsc{E-If} gives $l' \in H$ and $l' \guards t$.  Applying
    Lemma~\ref{lem:projection} to the initial premises of \textsc{E-If} and using
    the result and applying \textsc{E-If} gives $\typeRule{pc \sqcup l}{e_i}$.
    We can then apply \textsc{E-Bracket} to get the goal.
  \end{case}

@@

The noninterference theorem shows that if we have an expression that is typed
with a low label, then replacing any bracketed value in the expression with
either projection will result in the same value for the overall expression.  In
other words, if we vary any high security item in a low security expression, the
value must not be affected.

@@

We found that the process of mechanizing the lemmas that we proved was a useful
exercise in checking our understanding of our language, statements, and the work
of Simonet and Pottier which we based our approach off of.  We found ourselves
often surprised by the number of items that were elided by the proofs in the
paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
natural language but turned out to be 40 lines of Coq code, not including all of
the supporting definitions and rules.

@@

Once a full Coq proof of noninterference exists, a next step would be to
integrate this proof and all of the related definitions into a verified type
checking program, which could be integrated into a verified compiler.  We
already know that a full verified compiler has been done for C, and this gives
us hope that it would be possible to implement something similar for a language
like Jif.
